---
title: "Rcpp Part 2"
author: "Wenyu Wang"
date: "February 22, 2016"
output: html_document
---

```{r init, echo=FALSE}
library(Rcpp)
library(inline)
```

## Chapter 3. Data Structures

**RObject** is at the heart of Rcpp. 

Two core vector types "NumericVector" and "IntegerVector".

An instance of the RObject class encapsulates an R object. Every R object itself is internally represented by a SEXP: a pointer to a so-called S expression object.

### Integer vector class

```{r returnPerfectNumber}
src <- '
Rcpp::IntegerVector epn(4);
epn[0] = 6;
epn[1] = 14;
epn[2] = 496;
epn[3] = 8182;
return epn;
'

fun <- cxxfunction(signature(), src, plugin="Rcpp")
fun()
```

```{r useInput}
src <- '
Rcpp::IntegerVector vec(vx);
int prod = 1;
for (int i = 0; i < vec.size(); i++) {
  prod *= vec[i];
}
return Rcpp::wrap(prod);
'

fun <- cxxfunction(signature(vx = "integer"), src,
                   plugin="Rcpp")
fun(1:10)
```

Only the pointer to the underlying data is copied.

```{r useInput2}
src <- '
Rcpp::IntegerVector vec(vx);
int prod = std::accumulate(vec.begin(), vec.end(), 
                           1, std::multiplies<int>());
return Rcpp::wrap(prod);
'
fun <- cxxfunction(signature(vx = "integer"), src,
                   plugin = "Rcpp")
fun(1:10)
```

### Numeric vector class

```{r sumOfPower}
src <- '
Rcpp::NumericVector vec(vx);
double p = Rcpp::as<double>(dd);
double sum = 0.0;
for (int i = 0; i < vec.size(); i++) {
  sum += pow(vec[i], p);
}
return Rcpp::wrap(sum);
'

fun <- cxxfunction(signature(vx = "numeric", dd = "numeric"), src,
                   plugin = "Rcpp")
fun(1:4, 2.2)
```

```{r introduceClone}
src <- '
Rcpp::NumericVector invec(vx);
Rcpp::NumericVector outvec(vx);
for (int i = 0; i < invec.size(); i++){
  outvec[i] = log(invec[i]);
}
return outvec;
'

fun <- cxxfunction(signature(vx = "numeric"), src,
                   plugin = "Rcpp")
x <- c(1, 10, 100, 1000)
cbind(x, fun(x))
```

C++ object contains a pointer to the underlying SEXP object from R.

```{r introduceClone2}
src <- '
Rcpp::NumericVector invec(vx);
Rcpp::NumericVector outvec = Rcpp::clone(vx);
for (int i = 0; i < invec.size(); i++){
  outvec[i] = log(invec[i]);
}
return outvec;
'

fun <- cxxfunction(signature(vx = "numeric"), src,
                   plugin = "Rcpp")
x <- c(1, 10, 100, 1000)
cbind(x, fun(x))
```

This is better. Clone allocates memory for a new object.

```{r introduceClone3}
src <- '
Rcpp::NumericVector invec(vx);
Rcpp::NumericVector outvec = log(vx);
return outvec;
'

fun <- cxxfunction(signature(vx = "numeric"), src,
                   plugin = "Rcpp")
x <- c(1, 10, 100, 1000)
cbind(x, fun(x))
```

### Matrix

```{r matrix}
src <- '
Rcpp::NumericMatrix mat = 
  Rcpp::clone<Rcpp::NumericMatrix>(mx);
std::transform(mat.begin(), mat.end(), 
  mat.begin(), ::sqrt);
return mat;
'
fun <- cxxfunction(signature(mx = "numeric"), src,
                   plugin = "Rcpp")
mx <- matrix(1:9, nrow = 3)
fun(mx)
```

### Logical vector

```{r logicalvector}
fun <- cxxfunction(signature(), plugin="Rcpp",
                   body = '
Rcpp::LogicalVector v(6);
v[0] = v[1] = false;
v[2] = true;
v[3] = R_NaN;
v[4] = R_PosInf;
v[5] = NA_REAL;
return v;
                   ')
fun()
identical(fun(), c(F, F, T, NA, NA, NA))
```

## Chapter 4. Data structures

```{r someVec}
someVec <- c(mean=1.1, dim=10, cnt=10)
someVec
names(someVec)
```

Do the same thing in Rcpp.

```{r someVecRcpp}
src <- '
Rcpp::NumericVector x =
  Rcpp::NumericVector::create(
    Rcpp::Named("mean") = 1.1,
    Rcpp::Named("var") = 10,
    Rcpp::Named("cnt") = 20
  );
return x;
'

fun <- cxxfunction(signature(), src, plugin="Rcpp")
fun()
```

A much more concise version, 
1. using namespace Rcpp. 
2. _["key"] shortcut.

```{r someVecRcpp2}
src <- '
NumericVector x =
  NumericVector::create(
    _["mean"] = 1.1,
    _["var"] = 10,
    _["cnt"] = 20
  );
return x;
'

fun <- cxxfunction(signature(), src, plugin = "Rcpp")
fun()
```

### Generic vector class / List

